Qualifying Exam\phantom{:} \,\normalsize Physics 2020 @ YMSC, Tsinghua

### 2020-10-20 alpha/Exam/exam.tex
Rename submodule -> TeXplates & add tweaks
Including:
- build.sh script
- exam.tex

\begin{enumerate}
\item Event horizon(s): $f(r) = 0$, we have:
	\begin{enumerate}
	\item $M > \abs{Q}$, $r_\pm = M \pm \sqrt{M^2 - Q^2}$, 2 event horizons;
	\item $M = \abs{Q}$, $r_\pm = M$, 1 event horizons;
	\item $M < \abs{Q}$, no event horizons! ``Naked'' singularity.
	\end{enumerate}

@@

	\end{enumerate}
\end{enumerate}
\section{QFT}
We shall restore the reasonable convention: $
	\eta_{\mu\nu}
	\sim \pqty{-,+,+,+}
$. 
\begin{enumerate}
\item 1PI: diagrammatic contribution to the (1-particle) propagator that cannot be split into 2 disconnected parts by cutting one line; e.g.\,%
	\raisebox{-.4\baselineskip}{
		\feynmandiagram[
			scale=.4
			,small
			,layered layout
			,horizontal=a to d
		]{
			a -- b[{small,dot}] -- c[{small,dot}] -- d,
			b -- [half left,looseness=1.5]
			c -- [half left,looseness=1.5]
			b,
		};
	}
\item Consider the following Lagrangian:
	\begin{equation}
		\mcal{L}
		= -\frac{1}{2} Z(\pd\phi_r)^2
			-\frac{1}{2} m^2 Z\phi_r^2
			-\frac{\lambda}{4!} \phi_r^4
		-\frac{1}{2} \delta_Z(\pd\phi_r)^2
			-\frac{1}{2} \delta_m \phi_r^2
			-\frac{\delta_\lambda}{4!} \phi_r^4
	\label{eq:lagrangian}
	\end{equation}
	The convention here is rather bizarre; normally we write down the UV Lagrangian $\mcal{L}_{\mrm{UV}}$ and split it into 2 parts, one is the effective IR Lagrangian $\mcal{L}_{\mrm{IR}}$ and the other one is the counterterm:
	\begin{equation}
	\begin{aligned}
		\mcal{L}_{\mrm{UV}}
		&= -\frac{1}{2} Z(\pd\phi_r)^2
			-\frac{1}{2} m^2 Z\phi_r^2
			-\frac{\lambda}{4!} \phi_r^4 \\
		&= \pqty{
			-\frac{1}{2} (\pd\phi_r)^2
			-\frac{1}{2} m_p^2 \phi_r^2
			-\frac{\lambda_p}{4!} \phi_r^4
		}
		- \pqty{
			-\frac{1}{2} \delta_Z(\pd\phi_r)^2
			-\frac{1}{2} \delta_m \phi_r^2
			-\frac{\delta_\lambda}{4!} \phi_r^4
		} \\
		&= \mcal{L}_{\mrm{IR}}
			+ \mcal{L}_{\mrm{ct}}
	\end{aligned}
	\end{equation}

@@

	The $t$-integral is related to Beta functions; consider $t\mapsto \frac{t^2}{1 + t^2}$, and we have:
	\begin{equation}
		\int_0^\infty \frac{t^d \dd{t}}{1 + t^2}
		= \frac{1}{2}
		\int_0^1 t^{\frac{D}{2} - 1}
			(1-t)^{-\frac{D}{2}} \dd{t}
		= \frac{
				\Gamma(\tfrac{D}{2})\,
				\Gamma(1 - \tfrac{D}{2})
			}{2\Gamma(1)}
		= \frac{1}{2} \Gamma(\tfrac{D}{2})\,
			\Gamma(1 - \tfrac{D}{2})
		= \frac{\pi}{2\sin \frac{\pi D}{2}}
	\end{equation}
	The last line is the \textit{Euler's reflection formula}, but here we actually don't need that since the $\Gamma(\tfrac{D}{2})$ factor is canceled by $A(S^d)$. In the end we have:
	\begin{equation}
		\int \frac{\dd[D]{k}}{(2\pi)^D}
			\frac{1}{k^2 + m^2}
		= \frac{\pi^{D/2}}{(2\pi)^D}
			\Gamma(1 - \tfrac{D}{2})\,m^{D-2}
		= \frac{1}{(4\pi)^{D/2}}
			\Gamma(1 - \tfrac{D}{2})\,m^{D-2},
	\end{equation}
	\vspace{-.5\baselineskip}
	\begin{equation}
	%% ref: <https://tex.stackexchange.com/q/387810/193356>
		\raisebox{-.2\height}{
			\feynmandiagram[
				scale=.7
				,small
				,layered layout
				,horizontal=a to b
			]{
				a -- b [dot]
				-- [out=135, in=45, loop
					,min distance=3.5\baselineskip
					] b
				-- c
			};
		}
		= \frac{-i\lambda}{2} \frac{1}{(4\pi)^{D/2}}
			\Gamma(1 - \tfrac{D}{2})\,m^{D-2}
	\end{equation}

@@

	Alternatively, if we are working in $D = 4 = 3 + 1$ dimensions, it's easier to impose a naïve cutoff $\Lambda$, which gives:
	\begin{equation}
	\begin{aligned}
		\int^\Lambda \frac{k^d \dd{k}}{k^2 + m^2}
		&\sim \int^\Lambda k^{d-2} \dd{k}
			+ \int^\Lambda k^d \dd{k}
				\pqty{
					\frac{1}{k^2 + m^2}
					- \frac{1}{k^2}
				} \\
		&= \int^\Lambda k^{d-2} \dd{k}
			- m^2 \int^\Lambda 
				\frac{k^{d-2} \dd{k}}{k^2 + m^2},
		\quad d = D - 1 = 3 \\
		&= \frac{\Lambda^2}{2}
			- \frac{m^2}{2} \ln \pqty{
				1 + \frac{\Lambda^2}{m^2}
			},
	\end{aligned}
	\end{equation}
	Similarly, with $A(S^3) = 2\pi^2$, we have:
	\begin{equation}
	\begin{aligned}
		\delta_Z = 0,
	\quad
		\delta_m
		&= \frac{\lambda}{2}
			\frac{2\pi^2}{(2\pi)^4} \Bqty{
				\frac{\Lambda^2}{2}
				- \frac{m^2}{2} \ln \pqty{
					1 + \frac{\Lambda^2}{m^2}
				}
			} \\
		&= \frac{\lambda}{32\pi^2} \Bqty{
				\frac{\Lambda^2}{2}
				- \frac{m^2}{2} \ln \pqty{
					1 + \frac{\Lambda^2}{m^2}
				}
			}
	\end{aligned}
	\end{equation}

>>>>

\begin{enumerate}
\item Event horizon(s): $f(r) = 0$, we have:
	\begin{enumerate}
	\item $M > \abs{Q}$, $r_\pm = M \pm \sqrt{M^2 - Q^2}$, 2 event horizons;
	\item $M = \abs{Q}$, $r_\pm = M$, 1 event horizon;
	\item $M < \abs{Q}$, no event horizon! ``Naked'' singularity.
	\end{enumerate}

@@

	\end{enumerate}
\end{enumerate}
\section{QFT}
We shall restore the reasonable convention: $
	\eta_{\mu\nu}
	\sim \pqty{-,+,+,+}
$. 
\begin{enumerate}
\item 1PI: diagrammatic correction to the (1-particle) propagator that cannot be split into 2 disconnected parts by cutting one line; e.g.\,%
	\raisebox{-.4\baselineskip}{
		\feynmandiagram[
			scale=.4
			,small
			,layered layout
			,horizontal=a to d
		]{
			a -- b[{small,dot}] -- c[{small,dot}] -- d,
			b -- [half left,looseness=1.5]
			c -- [half left,looseness=1.5]
			b,
		};
	}
\item Consider the following Lagrangian:
	\begin{equation}
		\mcal{L}
		= -\frac{1}{2} Z(\pd\phi_r)^2
			-\frac{1}{2} m^2 Z\phi_r^2
			-\frac{\lambda}{4!} \phi_r^4
		-\frac{1}{2} \delta_Z(\pd\phi_r)^2
			-\frac{1}{2} \delta_m \phi_r^2
			-\frac{\delta_\lambda}{4!} \phi_r^4
	\label{eq:lagrangian}
	\end{equation}
	The convention here is rather bizarre; normally we write down the UV Lagrangian $\mcal{L}_{\mrm{UV}}$ and split it into 2 parts, one is the effective IR Lagrangian $\mcal{L}_{\mrm{IR}}$ and the other one is the counterterm:
	\begin{equation}
	\begin{aligned}
		\mcal{L}_{\mrm{UV}}
		&= -\frac{1}{2} Z(\pd\phi_r)^2
			-\frac{1}{2} m^2 Z\phi_r^2
			-\frac{\lambda}{4!} \phi_r^4 \\
		&= \pqty{
			-\frac{1}{2} (\pd\phi_r)^2
			-\frac{1}{2} m_p^2 \phi_r^2
			-\frac{\lambda_p}{4!} \phi_r^4
		}
		- \pqty{
			-\frac{1}{2} \delta_Z(\pd\phi_r)^2
			-\frac{1}{2} \delta_m \phi_r^2
			-\frac{\delta_\lambda}{4!} \phi_r^4
		} \\
		&= \mcal{L}_{\mrm{IR}}
			+ \mcal{L}_{\mrm{ct}}
	\end{aligned}
	\end{equation}

@@

	The $t$-integral is related to Beta functions; consider $t\mapsto \frac{t^2}{1 + t^2}$, and we have:
	\begin{equation}
		\int_0^\infty \frac{t^d \dd{t}}{1 + t^2}
		= \frac{1}{2}
		\int_0^1 t^{\frac{D}{2} - 1}
			(1-t)^{-\frac{D}{2}} \dd{t}
		= \frac{
				\Gamma(\tfrac{D}{2})\,
				\Gamma(1 - \tfrac{D}{2})
			}{2\Gamma(1)}
		= \frac{1}{2} \Gamma(\tfrac{D}{2})\,
			\Gamma(1 - \tfrac{D}{2})
		= \frac{\pi}{2\sin \frac{\pi D}{2}}
	\end{equation}
	The last line is \textit{Euler's reflection formula}, but here we actually don't need that since the $\Gamma(\tfrac{D}{2})$ factor is canceled by $A(S^d)$. In the end we have:
	\begin{equation}
		\int \frac{\dd[D]{k}}{(2\pi)^D}
			\frac{1}{k^2 + m^2}
		= \frac{\pi^{D/2}}{(2\pi)^D}
			\Gamma(1 - \tfrac{D}{2})\,m^{D-2}
		= \frac{1}{(4\pi)^{D/2}}
			\Gamma(1 - \tfrac{D}{2})\,m^{D-2},
	\end{equation}
	\vspace{-.5\baselineskip}
	\begin{equation}
	%% ref: <https://tex.stackexchange.com/q/387810/193356>
		\raisebox{-.2\height}{
			\feynmandiagram[
				scale=.7
				,small
				,layered layout
				,horizontal=a to b
			]{
				a -- b [dot]
				-- [out=135, in=45, loop
					,min distance=3.5\baselineskip
					] b
				-- c
			};
		}
		= \frac{-i\lambda}{2} \frac{1}{(4\pi)^{D/2}}
			\Gamma(1 - \tfrac{D}{2})\,m^{D-2}
	\end{equation}

@@

	Alternatively, if we are working in $D = 4 = 3 + 1$ dimensions, it's easier to impose a naïve cutoff $\Lambda$, which gives:
	\begin{equation}
	\begin{aligned}
		\int^\Lambda \frac{k^d \dd{k}}{k^2 + m^2}
		&\sim \int^\Lambda k^{d-2} \dd{k}
			+ \int^\Lambda k^d \dd{k}
				\pqty{
					\frac{1}{k^2 + m^2}
					- \frac{1}{k^2}
				} \\
		&= \int^\Lambda k^{d-2} \dd{k}
			- m^2 \int^\Lambda 
				\frac{k^{d-2} \dd{k}}{k^2 + m^2},
		\quad d = D - 1 = 3 \\
		&= \frac{\Lambda^2}{2}
			- \frac{m^2}{2} \ln \pqty{
				1 + \frac{\Lambda^2}{m^2}
			},
	\end{aligned}
	\end{equation}
	Similarly, with $A(S^3) = 2\pi^2$, we have:
	\begin{equation}
	\begin{aligned}
		\delta_Z = 0,
	\quad
		\delta_m
		&= \frac{\lambda}{2}
			\frac{2\pi^2}{(2\pi)^4} \Bqty{
				\frac{\Lambda^2}{2}
				- \frac{m^2}{2} \ln \pqty{
					1 + \frac{\Lambda^2}{m^2}
				}
			} \\
		&= \frac{\lambda}{32\pi^2} \Bqty{
				\Lambda^2
				- m^2 \ln \pqty{
					1 + \frac{\Lambda^2}{m^2}
				}
			}
	\end{aligned}
	\end{equation}
